---
title: "effects_factorial"
author: "Rich"
date: "2024-05-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters in the most complex models from the AI minimal group experiment.

This file looks at the factorial or deviation coding model output.

## load the libraries that we will be using ## 

## install ##

only install pkgs not already installed in wrangle or model files.

```{r install-pkg}
## install.packages(c("tinytable", "here"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "here",
         "brms", "tidybayes", "bayesplot", "future", "parallel", "tinytable")

lapply(pkg, library, character.only = TRUE)
```

## table options ##

```{r}
options(tinytable_tt_digits = 2,
        tinytable_print_output = "markdown")
```

# section 1 - read in the models and data #

## model ##

```{r}
bd3.3 <- readRDS("models/bd3.3.rds")
```

## data ##

may not be necessary, if we are only using the model objects

```{r}
# diffd <- read_csv("data/diffd.csv") 
# head(diffd)
```

# section 2 - Look at the point and interval estimates #

## summarise point and interval estimates for all aspects of the model ##

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(bd3.3) %>% 
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(bd3.3, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# varying effects - SD (all participant variation effects, 3 in total)
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# participant variation (8 varying effects x 238 pid = 1904 in total)
vary_vals_pt <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "pid")) %>% 
  median_qi(value)
head(vary_vals_pt)
tail(vary_vals_pt)
```

# section 2 - Look at the point and interval estimates #

## summarise point and interval estimates for all aspects of the model ##

using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts 
tidy_fixed <- post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(everything()) %>% # move from wide to long
  mutate(term = if_else(str_detect(name, "_group:mat2v1"), "group*MDvsMIP/MJP",
                if_else(str_detect(name, "_group:mat3v1"), "group*FAVvsP",
                if_else(str_detect(name, "_mat2v1"), "MDvsMIP/MJP",
                if_else(str_detect(name, "_mat3v1"), "FAVvsP",
                if_else(str_detect(name, "_group"), "group", "intercept"))))),
         term = factor(term, levels=c("intercept", "group", "MDvsMIP/MJP", "FAVvsP"
                                      , "group*MDvsMIP/MJP", "group*FAVvsP")))
head(tidy_fixed)
tail(tidy_fixed)

check.labels <- tidy_fixed %>% 
  distinct(name, term)
check.labels
                      
# plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "Model bd3.3 coefficient plot for fixed effects (predictors)",
       x = NULL, y = NULL) +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none") +
  coord_cartesian(xlim =c(-4, 4)) +
  scale_x_continuous(breaks=seq(-4,4,1))
p_tidy_fixed

ggsave ("figures/bd3.3_fixef.jpeg",
        width = 6, height = 4)
```

## make a table of fixed effects ##

```{r}
fix_tab <- post_qi_b %>%
  mutate(term = if_else(str_detect(name, "_group:mat2v1"), "group*MDvsMIP/MJP",
                if_else(str_detect(name, "_group:mat3v1"), "group*FAVvsP",
                if_else(str_detect(name, "_mat2v1"), "MDvsMIP/MJP",
                if_else(str_detect(name, "_mat3v1"), "FAVvsP",
                if_else(str_detect(name, "_group"), "group", "intercept"))))),
         term = factor(term, levels=c("intercept", "group", "MDvsMIP/MJP", "FAVvsP",
                                      "group*MDvsMIP/MJP", "group*FAVvsP"))) %>% 
  arrange(term) %>% 
  select(term, value, .lower, .upper)
fix_tab

# ## make a table using tinytable
tt_fix_tab <- tt(fix_tab,
                 caption="Model bd fixed effects") %>% 
  theme_tt("striped")
tt_fix_tab

## save it as a png
save_tt(tt_fix_tab, here("tables", "bd_fixed.png"),
        overwrite = T)

## save it as a word doc
save_tt(tt_fix_tab, here("tables", "bd_fixed.docx"),
        overwrite = T)

## tinytable is not perfect if you are just using it to create tables in word or as images
## it is probably better suited to creating manuscripts in r markdown with embedded tables in the code.

## what about flextable? I have issues installing it, specifically 'gdtools' package.
```

## Visualise varying effects ##

using posterior samples and tidybayes.

## Variance at the group level (i.e., intercept / b across participants) ##

```{r}
## varying effects of interest, including participants
tidy_varying <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(term = if_else(str_detect(name, "_mat2v1"), "MDvsMIP/MJP",
                if_else(str_detect(name, "_mat3v1"), "FAVvsP", "intercept")),
         term = factor(term, levels=c("intercept", "MDvsMIP/MJP", "FAVvsP")))
head(tidy_varying)
tail(tidy_varying)
str(tidy_varying)

check.varying <- tidy_varying %>% 
  distinct(name, term)
check.varying

p_tidy_varying <- ggplot(tidy_varying, aes(x = value, y = fct_rev(term))) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying pid effects",
       x = NULL, y = "SD") +
  theme_bw() +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank())
p_tidy_varying

ggsave ("figures/bd3.3_varying.jpeg",
        width = 6, height = 4)
```

## Variance within group levels i.e., at participant level ##

## filter and wrangle varying effects within the group level ##

Note - this should be in a centred metric so that deviation from the group average (zero) can be visualised. All good, but interpreting intercepts in some cases can difficult because it is when all other factors are zero. It can therefore be a little difficult to interpret and they may not be completely symmetric around zero for that reason. 

use output from posterior summary and wrangle

```{r}
# participant level
vary_pid <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "mat2v1"), "MDvsMIP/MJP",
                if_else(str_detect(param, "mat3v1"), "FAVvsP", "intercept")),
         term = factor(term, levels=c("intercept", "MDvsMIP/MJP", "FAVvsP"))) %>% 
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:500, times=3), # 500 pts, 3 variables/model parameters
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pid)
tail(vary_pid)
str(vary_pid)

check.vary.pid <- vary_pid %>% 
  group_by(term) %>% 
  tally()
check.vary.pid
```

# plot varying effects via caterpillar plots #

```{r}
p_tidy_varyvals <- ggplot(vary_pid, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  theme_bw()+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant")+
  # theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_wrap(~term)
p_tidy_varyvals

ggsave ("figures/bd3.3_varyvals.jpeg",
        width = 6, height = 4)
```
