---
title: "effects_index"
author: "Rich"
date: "2024-05-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters in the most complex models from the AI minimal group experiment.

This file looks at the index model output.

## load the libraries that we will be using ## 

## install ##

only install pkgs not already installed in wrangle or model files.

```{r install-pkg}
## install.packages(c("tinytable", "here"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "here",
         "brms", "tidybayes", "bayesplot", "future", "parallel", "tinytable")

lapply(pkg, library, character.only = TRUE)
```

## table options ##

```{r}
options(tinytable_tt_digits = 2,
        tinytable_print_output = "markdown")
```

# section 1 - read in the models and data #

## model ##

```{r}
## human group data 
bi_h2 <- readRDS("models/bi_h2.rds")

## ai group data
bi_ai2 <- readRDS("models/bi_ai2.rds")
```

## data ##

may not be necessary, if we are only using the model objects

```{r}
# diff_h <- read_csv("data/diff_h.csv") 
# head(diff_h)

# diff_ai <- read_csv("data/diff_ai.csv") 
# head(diff_ai)
```

# section 2 - Look at the point and interval estimates #

## summarise point and interval estimates for all aspects of the model ##

using posterior_samples and tidybayes

```{r}
## human model
posth <- as_draws_df(bi_h2) %>% 
  select(-lp__) %>% 
  as_tibble() %>% 
  add_column(group = "human", .before = 1)

## ai model
postai <- as_draws_df(bi_ai2) %>% 
  select(-lp__) %>% 
  as_tibble() %>% 
  add_column(group = "ai", .before = 1)

## combine dataframes
post <- bind_rows(posth, postai) %>% 
  mutate(group = factor(group,
                        levels = c("human", "ai")))
head(post)
tail(post)

## summarise parameter estimates per model
## human
post_summary_h <- posterior_summary(bi_h2, robust = TRUE)
## ai
post_summary_ai <- posterior_summary(bi_ai2, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(group, starts_with("b_")) %>% 
  pivot_longer(-group) %>% 
  group_by(group, name) %>% 
  median_qi(value)
head(post_qi_b)

# varying effects - SD (all participant variation effects, 6 in total)
post_qi_sd <- post %>%
  select(group, starts_with("sd_")) %>% 
  pivot_longer(-group) %>% 
  group_by(group, name) %>% 
  median_qi(value)
head(post_qi_sd)

# participant variation (3 varying effects x 249 pid + 3 varying effects x 251 = 1500 in total)
vary_vals_pt <- post %>%
  select(group, starts_with("r_")) %>% 
  pivot_longer(-group,
               values_drop_na = TRUE) %>% 
  group_by(group, name) %>% 
  filter(str_detect(name, "pid")) %>%
  median_qi(value)
head(vary_vals_pt)
tail(vary_vals_pt)
```

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts 
tidy_fixed <- post %>% 
  select(group, starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(-group) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         matrix_name = if_else(str_detect(key, "FAVvsMJP"), "FAVvsMJP",
                       if_else(str_detect(key, "MDvsMIPDMJP"), "MDvsMIP/MJP", "FAVvsP")),
         matrix_name = factor(matrix_name, levels=c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP")))
head(tidy_fixed)
tail(tidy_fixed)

check.labels <- tidy_fixed %>% 
  distinct(key, group, matrix_name) 
check.labels

# plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = group, y = value, fill=group)) +  
  # geom_vline(yintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "Model bi2 coefficient plot for fixed effects (predictors)",
       x = NULL, y = "response") +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none") +
  # coord_cartesian(xlim =c(-1.5, 3)) +
  # scale_x_continuous(breaks=seq(-1.5,3,1)) +
  facet_wrap(~matrix_name)
p_tidy_fixed

ggsave ("figures/bi_fixef.jpeg",
        width = 6, height = 4)
```

## make a table of fixed effects ##

```{r}
fix_tab <- post_qi_b %>%
  mutate(matrix_name = if_else(str_detect(name, "FAVvsMJP"), "FAVvsMJP",
                       if_else(str_detect(name, "MDvsMIPDMJP"), "MDvsMIP/MJP", "FAVvsP")),
         matrix_name = factor(matrix_name, 
                              levels=c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP"))) %>%
  select(matrix_name, group, value, .lower, .upper) %>% 
  arrange(matrix_name, group) 
fix_tab

## use kable and kableExtra
# kab_fix_tab <- kbl(fix_tab,
#                caption = "Model bi fixed effects",
#                digits = 2) %>% 
#   kable_styling() %>% 
#   save_kable("/Users/rramsey/Dropbox/exps/eliane/ai/analysis/tables/kab_bi_fixed.png")
# kab_fix_tab

# ## make a table using tinytable
tt_fix_tab <- tt(fix_tab,
                 caption="Model bi fixed effects") %>% 
  theme_tt("striped")
tt_fix_tab

## save it as a png
save_tt(tt_fix_tab, here("tables", "bi_fixed.png"),
        overwrite = T)

## save it as a word doc
save_tt(tt_fix_tab, here("tables", "bi_fixed.docx"),
        overwrite = T)
```

## Visualise varying effects ##

using posterior samples and tidybayes.

## Variance at the group level (i.e., intercept / b across participants) ##


```{r}
## varying effects of interest, including participants
tidy_varying <- post %>% 
  select(group, contains("sd_")) %>% 
  pivot_longer(-group) %>% 
  mutate(matrix_name = if_else(str_detect(name, "FAVvsMJP"), "FAVvsMJP",
                       if_else(str_detect(name, "MDvsMIPDMJP"), "MDvsMIP/MJP", "FAVvsP")),
         matrix_name = factor(matrix_name, levels=c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP")))
head(tidy_varying)
tail(tidy_varying)
str(tidy_varying)

check.varying <- tidy_varying %>% 
  distinct(group, matrix_name)
check.varying

p_tidy_varying <- ggplot(tidy_varying, aes(x = group, y = value)) +  
  # geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying pid effects",
       x = NULL, y = "SD") +
  theme_bw() +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank())+
  # coord_cartesian(xlim =c(0, 1))+
  # scale_x_continuous(breaks=seq(0,1,0.1)) +
  facet_wrap(~matrix_name)
p_tidy_varying

ggsave("figures/bi_varying.jpeg",
        width = 6, height = 4)
```

## Variance within group levels i.e., at participant level ##

## filter and wrangle varying effects within the group level ##

Note - this should be in a centred metric so that deviation from the group average (zero) can be visualised. All good, but interpreting intercepts in some cases can difficult because it is when all other factors are zero. It can therefore be a little difficult to interpret and they may not be completely symmetric around zero for that reason. 

use output from posterior summary and wrangle

```{r}
# participant level
# human group
vary_pid_h <- post_summary_h %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>%
  add_column(group = "human", .before = 1) %>% 
  mutate(matrix_name = if_else(str_detect(param, "FAVvsMJP"), "FAVvsMJP",
                       if_else(str_detect(param, "MDvsMIPDMJP"), "MDvsMIP/MJP", "FAVvsP")),
         matrix_name = factor(matrix_name, levels=c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP"))) %>% 
  arrange(matrix_name, Estimate) %>% 
  mutate(rank = rep(1:249, times=3), # 249 pts, 3 variables/model parameters
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pid_h)
str(vary_pid_h)

## ai group
vary_pid_ai <- post_summary_ai %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>%
  add_column(group = "ai", .before = 1) %>% 
  mutate(matrix_name = if_else(str_detect(param, "FAVvsMJP"), "FAVvsMJP",
                       if_else(str_detect(param, "MDvsMIPDMJP"), "MDvsMIP/MJP", "FAVvsP")),
         matrix_name = factor(matrix_name, levels=c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP"))) %>% 
  arrange(matrix_name, Estimate) %>% 
  mutate(rank = rep(1:251, times=3), # 251 pts, 3 variables/model parameters
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pid_ai)
str(vary_pid_ai)

## bind them together
vary_pid <- bind_rows(vary_pid_h, vary_pid_ai) %>% 
  mutate(group = factor(group,
                        levels = c("human", "ai")))
head(vary_pid)
str(vary_pid)

check.vary.pid <- vary_pid %>% 
  distinct(group, matrix_name)
check.vary.pid
```

## plot varying effects via caterpillar plots ##

```{r}
p_tidy_varyvals <- ggplot(vary_pid, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  theme_bw()+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant")+
  # theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_grid(matrix_name~group)
p_tidy_varyvals

ggsave ("figures/bi_varyvals.jpeg",
        width = 6, height = 4)
```