---
title: "wrangle"
author: "Rich"
date: "2024-05-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file wrangles raw data from the AI minimal group experiment. 
It produces some summary data plots, saves out data files for modelling and further analysis in later scripts.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
# install.packages(c("tidyverse", "RColorBrewer", "patchwork", "ggridges"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "ggridges")

lapply(pkg, library, character.only = TRUE)
```

## adjust plot settings ##

```{r plot-settings}
## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)

## theme settings for ggplot
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18), 
          title = element_text(size = 18),
          legend.position = "bottom")
)
```

## read in previously wrangled and saved data ##

this can sometimes be useful if you want to plot stuff without re-wrangling first.
Uncomment to load in these files.

```{r}
# data <- read_csv("data/data.csv") %>%
#   mutate(group = factor(group, levels = c("Human", "AI")),
#          matrix_name = factor(matrix_name, levels = c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP")),
#          condition = factor(condition, levels = c("together", "opposed")))
# head(data)
# 
# diff <- read_csv("data/diff.csv") %>%
#   mutate(group = factor(group, levels = c("Human", "AI")),
#          matrix_name = factor(matrix_name, levels = c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP")))
# head(diff)
```

# section 1 - basic wrangling and re-shaping #

## read in the raw data ##

```{r}
raw <- read_csv("data/AIHumanExperiment_Experiment.csv")
head(raw)
str(raw)
summary(raw)
glimpse(raw)
```

## rename variables and create factors ##

rename and select variables.

Just focus on a few key variables for now.

```{r}
data <- raw %>%
  rename(pid = participant, group = aihuman, all_conditions = condition, 
         matrix_name = matrixtype, condition = togetheropposed, 
         response = finalanswer) %>% 
  select(-jatosid, -prolificid, -all_conditions, -matrixnumber, -statvar, -answer, -rt)
head(data)
```

recode variables and create factors (this is useful for plotting the data)

```{r}
# recode and create factors
data <- data %>% 
  mutate(matrix_name = if_else(matrix_name == "1", "FAVvsMJP",
                       if_else(matrix_name == "2", "MDvsMIP/MJP", "FAVvsP")),
         condition = if_else(condition == "1", "together", "opposed"),
         group = factor(group, levels = c("Human", "AI")),
         matrix_name = factor(matrix_name, levels = c("FAVvsMJP", "MDvsMIP/MJP", "FAVvsP")),
         condition = factor(condition, levels = c("together", "opposed")))
head(data)
```

## take a quick look ##

overall summary

```{r}
summary(data)
```

tally by group

```{r}
tally <- data %>% 
  group_by(group) %>% 
  distinct(pid) %>% 
  tally()
tally
```

# section 2 - create summary data for plotting #

## at the individual pid level ##

```{r}
summary_pid <- data %>% 
  group_by(pid, group, matrix_name, condition) %>% 
  summarise(n= n(),
            mean=mean(response), 
            sd=sd(response), 
            sem = sd/sqrt(n))
head(summary_pid)
```

## at the group level ##

```{r}
summary_group <- data %>% 
  group_by(group, matrix_name, condition) %>% 
  summarise(n= n()/2, ## /2 as it counts both observations
            mean=mean(response), 
            sd=sd(response), 
            sem = sd/sqrt(n))
head(summary_group)
```

# section 3 - some plots #

violin

```{r}
p3.1 <- ggplot(summary_pid, 
              aes(x=condition, y=mean, fill = group)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), alpha = 1, 
               colour = "darkgrey") +
   geom_violin(alpha = 0.7, position=pd2) +
   geom_errorbar(data=summary_group, 
                 aes(y=mean, ymin=mean-sem, ymax=mean+sem), width=.2, position=pd2) +
   # geom_line(data=summary_group, aes(group=group), position=pd2) +
   geom_point(data=summary_group, aes(y=mean), position=pd2, size =1) +
   scale_colour_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
   labs(x="", y = "response")+
   scale_y_continuous(breaks=seq(0,12,2), limits=c(0,12)) +
   facet_grid(~matrix_name)+
   ggtitle("response by condition")
p3.1

ggsave ("figures/violin.jpeg",
        width = 8, height = 6)
```

# section 4 - calculate pull scores #

## calculate a difference score (or pull score) for each matrix type ##

For each matrix type, calculate the response on opposed minus the response on together. 
Since the responses on each of the together/opposed matrices can be 0-12, the difference score can be between -12 and +12.

This is now one difference score per pid and per matrix_name, which means 3 datapoints per pid.

```{r}
diff <- summary_pid %>% 
  group_by(pid, group, matrix_name) %>% 
  pivot_wider(-c(n, sd, sem),
              names_from = "condition",
              values_from = "mean") %>% 
  mutate(pull = opposed - together) %>% 
  ungroup()
head(diff)
```

# section 5 - create summary data on pull scores #

## at the group level ##

```{r}
diff_group <- diff %>% 
  group_by(group, matrix_name) %>% 
  summarise(n=n(),
            group_pull=mean(pull), 
            group_sd=sd(pull), 
            group_sem = group_sd / sqrt(n)) %>% 
  rename(pull = group_pull, sd = group_sd, sem = group_sem)
head(diff_group)
```

# section 6 - plot the pull scores #

now plot using violin plots

```{r}
p6.1 <- ggplot(diff, 
              aes(x=group, y=pull, fill = group)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), alpha = 1, 
               colour = "darkgrey") +
   geom_violin(alpha = 0.7, position=pd2) +
   geom_errorbar(data=diff_group, aes(y=pull, ymin=pull-sem, ymax=pull+sem), 
                 width=.2, position=pd2) +
   geom_line(data=diff_group, aes(y=pull, group=1), position=pd2) +
   geom_point(data=diff_group, aes(y=pull), position=pd2, size =1) +
   scale_colour_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
   labs(y = "response")+
   scale_y_continuous(breaks=seq(-12,12,2), limits=c(-12,12)) +
   guides(fill="none")+
   facet_grid(~matrix_name)+
   ggtitle("pull score by group and condition")
p6.1

ggsave ("figures/violin_diff.jpeg",
        width = 8, height = 6)
```

# section 7 - add deviation coding variables ##

create a new dataframe that we will use for the main analysis. 

```{r}
diffd <- diff %>% 
  mutate(groupd = if_else(group == "AI", 0.5, -0.5),
         mat2v1 = if_else(matrix_name == "MDvsMIP/MJP", 2/3, -1/3), # target MDvsMIP/MJP
         mat3v1 = if_else(matrix_name == "FAVvsP", 2/3, -1/3)) %>%  # target FAVvsP
  select(pid, groupd, mat2v1, mat3v1, pull) %>% 
  rename(group=groupd)
head(diffd)
```

# section 8 - save out some data #

```{r}
write_csv(data, "data/data.csv") # all data good for plotting
write_csv(summary_pid, "data/summary_pid.csv")
write_csv(summary_group, "data/summary_group.csv")
write_csv(diff, "data/diff.csv")
write_csv(diff_group, "data/diff_group.csv")
write_csv(diffd, "data/diffd.csv")
```


# section 9 - play around with some density plots #

density (raw scores)

```{r}
p9.1 <- ggplot(data, 
              aes(x=response, fill = group)) +
   geom_density(alpha=0.5) +
   facet_grid(matrix_name~condition) +
   ggtitle("response by group, matrix type and condition")
p9.1
```

density (pull scores)

```{r}
p9.2 <- ggplot(diff, 
              aes(x=pull, fill = group)) +
   geom_density(alpha=0.5) +
   facet_grid(~matrix_name) +
   ggtitle("pull score by group and matrix type")
p9.2
```

density (log(pull) scores)

first, create a log(pull)

```{r}
diff <- diff %>% 
  mutate(pos_pull = pull + 13,
         log_pull = log(pos_pull))
head(diff)
summary(diff)
```

plot positive pull scores

```{r}
p9.3 <- ggplot(diff, 
              aes(x=pos_pull, fill = group)) +
   geom_density(alpha=0.5) +
   facet_grid(~matrix_name) +
   ggtitle("log(pull) score by group and matrix type")
p9.3
```

plot log pull scores

```{r}
p9.4 <- ggplot(diff, 
              aes(x=log_pull, fill = group)) +
   geom_density(alpha=0.5) +
   facet_grid(~matrix_name) +
   ggtitle("log(pull) score by group and matrix type") +
   xlim(0,4)
p9.4
```

plot together

```{r}
p9.5 <- p9.2 / p9.4 
p9.5
```

let's try to display percentiles for the size of the difference score

using ggridges and with data points too

```{r plot9.6}
p9.6 <- ggplot(diff, aes(x = pull, y = "", fill = after_stat(quantile))) +
  stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 10, quantile_lines = TRUE, 
    jittered_points = TRUE, scale = 10,
    linewidth = 0.7, vline_color = "red",
    point_size = 1, point_alpha = 0.25, 
    position = position_raincloud(adjust_vlines = TRUE)
  ) +
  scale_fill_viridis_d(name = "Quantiles") +
  geom_vline(xintercept = 0, colour = "black", linetype = "dashed",
             linewidth = 1) +
  geom_vline(data = diff_group,
             aes(xintercept = pull), colour = "turquoise", linetype = "dashed",
             linewidth = 1) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1)) +
  labs(title = "Pull score density quantiles by group and matrix type",
       y = NULL,
       x = "pull score") +
  # scale_x_continuous(breaks = seq(-150, 150, 25)) +
  facet_grid(group~matrix_name)
p9.6

ggsave("figures/diff_dens_q.jpeg",
       width = 8, height = 6)
```